### 2024-09-09 Batching Draft 8/Batching Draft 8.Rmd
getting going

---
title: "To Batch or Not to Batch: Test-Ordering Variability in the ED"
authors:
  - name: Jacob Jameson
    department:  Interfaculty Initiative in Health Policy
    affiliation: Harvard University
    location: Cambridge, MA 
  - name: Soroush Saghafian
    department:  Harvard Kennedy School of Government
    affiliation: Harvard University
    location: Cambridge, MA 
  - name: Robert Huckman
    department:  Harvard Business School
    affiliation: Harvard University
    location: Boston, MA 
  - name: Nicole Hodgson
    department:  Emergency Department
    affiliation: Mayo Clinic
    location: Phoenix, AZ
abstract: |

keywords:

bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
header-includes:
  - \usepackage{booktabs}
  - \usepackage{tabularx}
  - \usepackage{graphicx}
  - \usepackage{tikz}
  - \usepackage{siunitx}
  - \usepackage{tablefootnote}
  - \usepackage{longtable}
  - \usepackage{threeparttable}
  - \usepackage{natbib}
  - \usepackage{caption}
  - \usepackage{adjustbox}
  - \usepackage{multirow}
---

# Introduction

Healthcare delivery, particularly in the emergency department (ED), is a delicate balance that involves ensuring optimal patient outcomes while optimizing resource utilization. Achieving these twin goals requires timely and accurate diagnosis, which in turn enables prompt and appropriate treatment, consequently improving patient prognosis and reducing the likelihood of adverse events. Furthermore, efficient patient discharge from the ED can help alleviate overcrowding, a severe issue with potential consequences including higher complication rates and increased mortality [@bernstein2009].

One important factor that can impact the speed and effectiveness of diagnosis in the ED is the availability and performance of diagnostic tests [@naseim2015]. A variety of diagnostic tests are used in the ED, including laboratory tests, imaging studies, and specialized tests. These tests can provide valuable information about a patient’s condition and help to guide treatment decisions.

A critical question in this context pertains to whether physicians in the ED should batch order diagnostic tests or order them sequentially. This decision essentially represents a tradeoff between reducing patient length of stay and risk of over-testing. Over-testing, or performing unnecessary tests, can lead to increased costs, unnecessary patient anxiety, and potential harm from follow-up of false-positive results [@koch2018]. Conversely, keeping a patient for an extended time to perform all possible tests could lead to ED overcrowding, an issue associated with severe consequences, as mentioned earlier. Instead, what is needed is a reasonable balance between the number of diagnostic tests performed and the total time the patient is kept in the ED before either being admitted or discharged. Several studies have demonstrated that optimizing the ED patient flow process can result in significant improvements [@saghafian2015], however, research surrounding test ordering strategies to improve the patient flow processes remains limited.

In this paper, we use data from two large EDs to quantify the benefits and consequences of batching versus sequentially ordering advanced imaging tests on patient length of stay, re-admission, and total imaging volume. Our empirical strategy exploits random assignment of patients to ED physicians who differ in their propensity to batch-order diagnostic tests. When patients arrive at the ED, they are assigned to a physician based on availability, with no discretion on either side. Thus, patients who arrive at the ED at similar times are randomly assigned to physicians who vary in their willingness to batch order diagnostic tests. We measure physician tendency to batch using a leave-out, residualized measure based on all other patients the physician has seen in the ED in the study period. The tendency measure strongly predicts the ED test batch outcome but is uncorrelated with patient and ED visit characteristics.

With the caveat that other unobserved dimensions of physician care may impact patient outcomes, we employ physician batch tendency as an instrumental variable for having tests batch ordered in the ED to quantify the effect of batching directly. Because of the institutional features of the ED, our research design closely approximates an RCT that assigns patients to batch-ordering or sequential-ordering arm. In the ED, patients have no discretion over choosing providers, and in our specific ED, physicians have discretion over choosing patients, alleviating major selection issues present in other health care settings. Furthermore, physicians exhibit wide variation in practice behavior in batch-ordering, even within the same hospital, while following the same guidelines. Finally, patient-physician interactions in the ED are typically well documented, short, and one-off, constraining physician decision-making to a more limited, better-observed choice set than present in settings such as specialty or primary care.

In sum, exploiting practice variation in ED settings shuts down other (but not all) potential channels besides test batching that are present in other settings, determine length of stay, and impact patient outcomes. This approach allows us to move closer to identifying the causal impact of batch-ordering diagnostic tests on patient outcomes and resource utilization. It is important to note that this paper studies the impact of batch-ordering through a batching decision requiring clinical judgment (within practice norms) rather than through specific hospital policies, differences in adherence to clinical practice guidelines, or substandard care.

# Literature Review

# Theoretical Background and Hypotheses

Emergency departments (EDs) are complex, time-sensitive environments where physicians must make rapid decisions under uncertainty [@asplin2003model; @jarvis2016ed]. Our theoretical model aims to capture the nuanced dynamics of diagnostic decision-making in this high-stakes setting, with a particular focus on the practice of batch ordering imaging tests.

## Diagnostic Decision-Making and Test Ordering in the ED

In the ED, physicians face the challenge of diagnosing and treating patients efficiently while maintaining high-quality care [@croskerry2013cognitive; @johnson2019decision]. Upon a patient's arrival, the physician must decide on a diagnostic strategy, which often involves ordering imaging tests. This decision-making process is influenced by various factors, including the physician's experience, the patient's presenting symptoms, and the ED's current state (e.g., crowding levels) [@cooke2013overview; @bernstein2009effect; @morley2018emergency].

The key distinction in our study is between batch ordering and sequential ordering of imaging tests. In batch ordering, a physician orders multiple tests simultaneously, while in sequential ordering, tests are ordered one at a time, with each subsequent test decision informed by the results of the previous tests. It's crucial to understand that batch ordering does not mean tests are performed simultaneously. Instead, it means that patients enter the queues for multiple tests at once. This distinction is important when considering the implications of batch ordering on ED operations and patient flow [@green2006managing; @batt2019impact].

## Queuing Theory and ED Operations

When a physician batch orders tests, the patient enters multiple queues simultaneously. This can be modeled using parallel queuing systems in operations research [@dai2010load; @zychlinski2019discrete]. While this might seem to expedite the process, it can lead to inefficiencies in resource allocation, as multiple departments (e.g., radiology, laboratory) must allocate resources to a single patient simultaneously, potentially leading to suboptimal resource utilization [@shi2016models; @wang2019data].

Moreover, batch ordering creates an information lag where results from one test cannot inform the decision to proceed with subsequent tests, potentially leading to unnecessary testing [@welch2011overdiagnosed; @o2018overdiagnosis]. This approach can also create blocking effects, where a patient waiting for multiple tests may block other patients who need only one of those tests, potentially increasing overall waiting times [@armony2015heterogeneous; @song2020managing].

## Reasons for Batch Ordering and Theoretical Implications

Physicians may choose to batch order for several reasons. They may believe it will speed up the diagnostic process [@kocher2012effect; @rosen2018quality], or they may be practicing defensive medicine to avoid missing critical diagnoses [@studdert2005defensive; @kanzaria2015emergency]. Batch ordering may also reduce the cognitive load required to make multiple sequential decisions [@croskerry2013cognitive; @nibbelink2018emergency], which can be particularly appealing in busy EDs where physicians are under time pressure [@bernstein2009effect; @morley2018emergency].

From a theoretical perspective, batch ordering presents several implications. It forgoes the opportunity for value of information considerations, where each test result could inform subsequent decisions [@smith2013impact; @akhlaghpour2020value]. Sequential ordering preserves the option to stop testing once a diagnosis is reached, aligning with real options theory in decision-making under uncertainty [@dixit1994investment; @garg2018value]. While batch ordering might seem efficient for individual patients, it could lead to systemic inefficiencies in resource utilization and patient flow [@green2006managing; @batt2019impact].

## Hypotheses Development

Based on this theoretical framework, we develop the following hypotheses:

### Impact on Resource Utilization

*Hypothesis 1: Batch ordering of diagnostic tests will lead to increased overall test utilization compared to sequential ordering.* This hypothesis is grounded in the idea that batch ordering bypasses the opportunity to use information from initial tests to inform the necessity of subsequent tests, potentially leading to unnecessary testing [@welch2011overdiagnosed; @o2018overdiagnosis].

### Effect on Length of Stay

*Hypothesis 2: Batch ordering will have a significant impact on patient length of stay (LOS) in the ED, but the direction of this effect is ambiguous.* While batch ordering might reduce LOS by eliminating waiting times between test decisions [@kocher2012effect; @rosen2018quality], it could also increase LOS due to potential queuing inefficiencies and the time required to perform and interpret additional tests [@forster2003effect; @armony2015heterogeneous; @song2020managing].

### Quality of Care and Return Visits

*Hypothesis 3: Batch ordering will influence the rate of 72-hour return visits with admission, but the direction of this effect is uncertain.* Comprehensive initial testing might reduce missed diagnoses and thus return visits [@pines2010national; @sabbatini2016overzealous]. However, it could also lead to false positives and unnecessary follow-ups, potentially increasing return visits [@welch2011overdiagnosed; @o2018overdiagnosis].

### Hypothesis 4: Heterogeneity Across Presenting Complaints

*Hypothesis 4: The effects of batch ordering on resource utilization, length of stay, and quality of care will vary significantly across different presenting complaints.* This hypothesis is based on the understanding that different clinical presentations have varying levels of diagnostic uncertainty and risk [@cooke2013overview; @johnson2019decision], which may influence the appropriateness and impact of batch ordering.

By testing these hypotheses, we aim to provide a comprehensive understanding of the impacts of batch ordering in the ED, informing both clinical practice and operational policy. Our empirical strategy, detailed in the following sections, is designed to rigorously test these hypotheses and uncover the causal effects of batch ordering on ED performance and patient outcomes.

# Data and Methods

Our study uses data from two large U.S. emergency departments (EDs): the Mayo Clinic of Arizona and Massachusetts General Hospital (MGH). The MGH dataset, which includes 129,489 patient encounters from November 10, 2021 through December 10, 2022, provides a robust sample for validating the generalizability of our findings. However, our primary analysis focuses on the Mayo Clinic data due to its unique feature of random patient-physician assignment, which allows for stronger causal inference.

## Mayo Clinic Data

Our primary dataset comes from the ED of the Mayo Clinic of Arizona, a tertiary care hospital without obstetrical services, an inpatient pediatrics unit, or a trauma designation. During the study period from October 6, 2018, through December 31, 2019, the ED recorded 48,854 visits per year, managed across 26 treatment rooms and up to 9 hallway spaces. The department is exclusively staffed by board-eligible or board-certified emergency physicians (EPs), with rotating residents overseeing about 10% of patient volume. The data is summarized in Table \ref{tab:summary_stats}.

\begin{table}[ht]
\centering
\caption{Summary Statistics of Mayo Clinic Emergency Department Encounters}
\label{tab:summary_stats}
\begin{threeparttable}
\begin{tabular}{p{10cm}ccccc}
\toprule
Variable &  & {Mean} & {Q1} & {Median} & {Q3} \\
\midrule
\textbf{Emergency Department Characteristics} & & & & & \\
Total Patients & & {48,854} & {-} & {-} & {-} \\
Patients Admitted & & {18.7\%} & {-} & {-} & {-} \\
Patients Revisited within 72 Hours & & {3.81\%} & {-} & {-} & {-} \\
Patients with IV Fluids & & {35.7\%} & {-} & {-} & {-} \\
Patients with IV Meds & & {17.9\%} & {-} & {-} & {-} \\
Patients in ED & & {24.1} & {-} & {-} & {-} \\
Tachycardic & & {19.2\%} & {-} & {-} & {-} \\
Tachypneic & & {8.82\%} & {-} & {-} & {-} \\
Febrile & & {2.16\%} & {-} & {-} & {-} \\
Hypotensive & & {1.42\%} & {-} & {-} & {-} \\
ESI & & {2.81} & {-} & {-} & {-} \\
Time from Arrival to Triage (mins) & & {8.03} & {4} & {6} & {10} \\
Time from Triage to First Contact (mins) & & {80.2} & {11} & {29} & {61} \\
Average ED LOS (min) & & {246} & {-} & {-} & {-} \\
\midrule
\textbf{Patient Demographics} & & & & & \\
Percent Male & & {46.5\%} & {-} & {-} & {-} \\
Race: White & & {88.4\%} & {-} & {-} & {-} \\
Race: Black & & {4.17\%} & {-} & {-} & {-} \\
Race: Asian & & {3.05\%} & {-} & {-} & {-} \\
Gender: Female & & {53.5\%} & {-} & {-} & {-} \\
Arrival Age & & {57.7} & {43} & {61} & {74} \\
\midrule
\textbf{Diagnostic Tests and Outcomes} & & & & & \\
X-Rays Performed & & {43.3\%} & {-} & {-} & {-} \\
Ultrasounds Performed & & {11.3\%} & {-} & {-} & {-} \\
CTs Performed & & {35.5\%} & {-} & {-} & {-} \\
Labs Ordered & & {73.7\%} & {-} & {-} & {-} \\
Patients Discharged & & {66.8\%} & {-} & {-} & {-} \\
Patients Admitted & & {18.7\%} & {-} & {-} & {-} \\
Contrast CT Performed & & {17.7\%} & {-} & {-} & {-} \\
Time to Result: X-Ray (mins) & & {67.2} & {36} & {54} & {79} \\
Time to Result: Ultrasound (mins) & & {165} & {71} & {101} & {150} \\
Time to Result: Contrast CT (mins) & & {142} & {86} & {115} & {153} \\
Time to Result: Non-Contrast CT (mins) & & {89.7} & {50} & {70} & {102} \\
Time to Result: Lab (mins) & & {46.1} & {25} & {35} & {49} \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item \textit{This table reports summary statistics for the baseline sample of emergency department visits during the study period described in the text. Vital signs were categorized as follows: tachycardia (pulse more significant than $100$), tachypnea (respiratory rate greater than $20$), fever (temperature greater than $38^\circ C$), and hypotension (systolic blood pressure less than $90$).}
\end{tablenotes}
\end{threeparttable}
\end{table}

A key feature of the Mayo Clinic ED is its rotational patient assignment system, in which patients arriving at the Mayo Clinic ED are randomly assigned to physicians via a rotational patient assignment algorithm @traub2016emergency, which removes potential selection bias concerns for our analyses. In essence, barring arrival time and shift-level variation, the physician-to-patient matching can be deemed random. Table \ref{tab:wald_test} displays that patient encounters (regarding chief complaints and emergency severity) are equitably distributed across physicians within our study's cohort.

\begin{table}[htbp]
    \centering
    \caption{Balancing Test: Wald Test for Equality of Means}
    \label{tab:wald_test}
    \begin{tabular}{p{10cm}cccc}
        \toprule
        Chief Complaints & Frequency & F-Statistic & $Pr(> F)$ \\
        \midrule
        Abdominal Complaints & 6,232 & 2.587 & 0.108 \\ 
        Back or Flank Pain & 2,552 & 1.637 & 0.201 \\ 
        Chest Pain & 3,525 & 0.407 & 0.524 \\ 
        Extremity Complaints & 5,265 & 1.847 & 0.174 \\ 
        Falls, Motor Vehicle Crashes, Assaults, and Trauma & 2,381 & 0.023 & 0.880 \\ 
        Gastrointestinal Issues & 3,323 & 0.105 & 0.746 \\ 
        Neurological Issue & 3,495 & 0.135 & 0.713 \\ 
        Shortness of Breath & 2,966 & 1.324 & 0.250 \\ 
        Skin Complaints & 2,178 & 0.383 & 0.536 \\ 
        Upper Respiratory Symptoms & 1917 & 0.017 & 0.896 \\ 
        \midrule
        Emergency Severity & Frequency & F-Statistic & $Pr(> F)$ \\
        \midrule
        ESI 1 or 2 & 13,914 & 0.011 & 0.915 \\ 
        ESI 3, 4, or 5 & 29,386 & 0.010 & 0.921 \\ 
        \bottomrule
    \end{tabular}
\begin{tablenotes}
\small
\item \textit{Table \ref{tab:wald_test} reports the results of a Wald test which was conducted to assess the balance of chief complaints across providers in our dataset. A balanced distribution implies that complaints and severity are evenly distributed across providers, which we expect to be the case due to randomization. The Wald F-statistic and p-value are reported. Robust standard errors (type HC1) were used to account for potential heteroscedasticity in the data.}
\end{tablenotes}
\end{table}

We conducted a retrospective review of comprehensive ED operational data, coinciding with the initiation of a new electronic medical record. The dataset includes detailed patient demographics, chief complaints, vital signs, emergency severity index (ESI), length of stay (LOS), and resource utilization metrics. This period was chosen to provide a robust data set while excluding the influence of the coronavirus pandemic.

## Sample Construction

Our research design focuses on adults who visit the Mayo Clinic of Arizona ED. We observe $48,854$ such visits during the study period. To improve power, we drop encounters with rare chief complaints ($<1000$ total encounters of this kind) and complaints where a batch order occurs less than $5$ percent of the time. Since batch orders are rare for these cases, our physician batch tendency instrument could suffer from a weak instrument problem if we included them. Examples of complaints dropped include Upper Respiratory Symptoms and Urinary Complaints. Excluding these conditions does not introduce selection bias unless physician test batching tendency is orthogonal to physician diagnosing behavior. While this assumption may be violated if we were to use a very detailed level of chief complaint information upon which to base our exclusion criterion, it is plausibly satisfied when using broad complaint categories as we do. In order to estimate a precise measure of physician-level batch tendency, we further restrict our sample to the $15,821$ encounters involving full-time physicians who treat over 500 ED cases per year.

## Variable Definitions

Our explanatory variable in the IV analysis, $Batched_i$, is an indicator for whether patient $i$ has their tests batch-ordered at their ED encounter. While the patient could decide not to undergo the tests ordered by the physician, this is rare in practice. We define "batching" in line with standard emergency medicine practices. Batching occurs when a physician simultaneously orders a comprehensive set of diagnostic tests, typically covering a broad range of potential diagnoses. This contrasts with sequential ordering, where tests are ordered sequentially based on the information obtained from each test as needed.

We operationalize batching as occurring when multiple diagnostic imaging tests are ordered within a 5-minute window. In Section [...], sensitivity analyses on this cutoff point showed that our results are robust to this definition. Each imaging test (e.g., X-ray, CT scan, Ultrasound) is considered a separate, distinct test for our study. Therefore, a batch in our study consists of two or more distinct imaging tests.

Our dependent variables are as follows:

*(a) ED length of stay (LOS).*-- ED-LOS is a critical measure of efficiency and patient throughput in emergency care settings. It is defined as the duration from a patient's arrival to the ED until their departure, whether by discharge or admission to the hospital.

*(b) Number of Unique Imaging Tests*-- Resource utilization in the ED typically refers to the extent of medical services and interventions a patient receives. In this study, we quantify resource utilization by the total number of distinct imaging tests ordered per patient during their ED stay. This encompasses both initial and subsequent tests.

*(c) 72 Hour Return with Admission*-- This is a binary variable indicating whether the patient returns to the ED within 72 hours of their initial visit and is admitted to the hospital. This outcome is used to assess the quality of care provided during the initial ED visit.

## Empirical Strategy

Our empirical strategy closely follows the literature that relies on quasi-random assignment of agents to cases, often referred to as the “judges design.” Papers in this literature typically exploit variation in the sentencing leniency of judges who work in the same court. Similarly, we explore batching variation across physicians who work in the same emergency department. In its reduced form, under the assumption of quasi-random assignment, this approach allows researchers to identify the causal effect of being assigned to different types of physicians. Under additional assumptions, an instrumental variable approach identifies the causal effect of a given medical decision. We employ both approaches and lay out their details in the next subsections.

### Batch Tendency Construction

To measure physician batch tendency, we use the physician's residualized leave-out average batch rate. This measure is derived from two steps following the approaches taken by @doyle2015measuring, @dobbie2018effects, and @eichmeyer2022pathways. First, we obtain residuals from a regression model, which includes all ED encounters in our sample period.

\begin{equation}
Batched_{i,t} = \alpha_0 + \alpha_{ym} + \alpha_{dt} + \alpha_{complaint\_esi} + \alpha_{lab} + \varepsilon_{i,t}
\end{equation}

Where $Batched_{i,t}$ is a dummy variable equal to one if patient $i$ had their imaging tests batch ordered on encounter that took place on date $t$. Fixed effects include year-month fixed effects, $\alpha_{ym}$, to control for time and seasonal variation in batching, such as hospital-specific policies (e.g. initiatives to eliminate excess testing) or seasonality in ED visits. We also control for ``shift-level" variations that include both physician scheduling and patient arrival with day of week-time of day fixed effects, $\alpha_{dt}$. Chief complaint by severity fixed effects, $\alpha_{complaint}$, were also included to increase precision. Finally, a binary variable for whether or not laboratory tests were ordered, $\alpha_{lab}$, was included to account for the complexity of the case. As stated earlier, these controls are more than what is required for our quasi-random assignment assumption. Under the assumption that we have captured the observables under which quasi-random assignment occurs in the ED, the unexplained variation-- the physician’s contribution-- resides in the error term, $\varepsilon_{i,t}$.

In step two, the tendency measure for patient $i$ seen by physician $j$ is computed as the average residual across all other patients seen by the physician that year:

\begin{equation}
Tendency_{i,j}^{phys} =
\frac{1}{N_{-i,j}} \sum_{i' \in \{J \backslash i\}}\hat{\varepsilon}_{i'}
\end{equation}

where $\hat{\varepsilon}_{i'} = \hat{Batch}_{i'} - Batch_{i'}$ is the residual from equation (1); $J$ is the set of all ED encounters treated by physician $j$; and $N_{-i,j} = |\{J \backslash i\}|$, the number of cases that physician has seen that year, excluding patient $i$. This leave-out mean eliminates the mechanical bias that stems from patient $i$'s own case entering into the instrument. The measure is interpreted as the average (leave-out) batch rate of patient $i$’s physician, relative to other physicians in that hospital-year-month, hospital-day of week-time of day.

We document that the Mayo Clinic ED physicians exhibit wide, systematic variation in their propensity to batch order imaging tests. Table \ref{table:first_stage} presents the ``first stage" in a regression table: being assigned to a $10$pp higher batch-tendency physician is associated with a $6$pp increase in the likelihood of having tests batch-ordered in the ED. The F-statistic is 94 when all controls and fixed effects are included. The coefficient is greater than one because all emergency visits are used to construct the tendency instrument, while the first stage is calculated using the baseline sample only, which excludes the rare complaints.

\begin{table}[!htbp] \centering 
  \caption{Comparison of First Stage Estimates}
  \label{table:first_stage}
  \begin{tabularx}{5.5in}{Xccc} % 'X' for the first column and 'c' for the others
  \toprule
   & \multicolumn{3}{c}{Coeffecient} \\
  \cmidrule{2-4}
   & (1) & (2) & (3) \\
  \midrule
  Batch Tendency & 1.92$^{***}$ & 1.91$^{***}$ & 1.76$^{***}$ \\ 
   & (0.07) & (0.06) & (1.76) \\ 
  \textit{Day of Week-Time of Day FE} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
  \textit{Month of Year FE} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
  \textit{Complaint/Severity FE} & & $\checkmark$ & $\checkmark$ \\
  \textit{Laboratory Tests Ordered} & & & $\checkmark$ \\
  \midrule
  F Statistic & 9.55 & 17.86 & 21.59 \\ 
  N & 16,361 & 16,361 & 16,361 \\ 
  \bottomrule
  \end{tabularx}
  \begin{tablenotes}
  \small
  \item Estimates of the first stage for the baseline sample described in the text. Seasonality shift fixed effects include Year-Month and Hospital-Day of week-Hour of day fixed effects. Chief complaint comes from the cleaned complaint that the patient came in with at the initial encounter. Column 3 corresponds to the baseline controls. Robust standard errors are clustered at the physician level.
  \item $^{***} p < 0.001$, $^{**} p < 0.01$, $^{*} p < 0.05$.
  \end{tablenotes}
\end{table}

To estimate the reduced-form effects of being treated by a batch-preferring physician, we estimate the following equation:

\begin{equation}
Y_i = \mu_0 + \mu_1 Tendency_{i,j}^{phys} + \gamma X_i + \nu_i
\end{equation}

This reduced form will allow us to check that our instrument is a strong instrument. To study the effects of test batching in the ED on an outcome $Y_i$, we estimate the following 2SLS equations using our baseline sample:

\begin{equation}
Y_i = \beta_0 + \beta_1 Batched_i + \theta X_i + \varepsilon_i
\end{equation}

\begin{equation}
Batched_i = \delta_0 + \delta_1 Tendency_{i,j}^{phys} + \delta_2 X_i + \nu_i
\end{equation}

Where $Y_i$ represents our main outcomes of interest: length of stay, 72 hour return with admission, and number of imaging tests ordered for the patient, and $X_i$ is the same as in the reduced-form approach. $Batched_i$ variable suffers from potential endogeneity concerns. For example, injury severity may be unobserved and correlated with need to run multiple tests, length of stay, and return with admission likelihood. Hence, we instrument $Batched_i$ with the assigned physician $j$'s underlying tendency to batch, $Tendency_{i,j}^{phys}$. We cluster robust standard errors at the physician level to account for the assignment process of patients to physicians.

## Identifying Assumptions

The reduced-form approach delivers an unbiased estimate of the causal effect of being treated by a higher tendency to batch physician, since assignment of patients to ED physicians is random, conditional on seasonality and shift (``conditional independence"). The residualization in equation (1) controls for more controls than required to achieve quasi-random assignment; they are included for statistical precision in measuring physician tendency to batch.

Our instrumental variable approach, which aims to recover the causal effect of having diagnostic tests batch ordered, relies on three additional assumptions: relevance, exclusion, and monotonicity. We reported a strong first stage (i.e., relevance) at the end of the previous Section. The exclusion restriction requires that the instrument must influence the outcome of interest only through its effect on test batching. This is perhaps our strongest assumption and is at its core, untestable. However, several features of the ED setting suggest that such violation may likely only have a small impact and may be less concerning than in other health care settings. First, unlike in primary care settings, where the patient and primary care provider have many repeat encounters, the scope of what the emergency physician can do to impact medium-term outcomes is limited and well-observed by the researcher. Second, any violation of the exclusion restriction needs to directly affect the specific outcome of interest. The channel by which ED physicians can influence length of stay relative outcomes is likely through testing and diagnosis. Nevertheless, we take this assumption seriously and perform a placebo check in Section [...] as well as various robustness checks in Section [...].

Finally, the monotonicity assumption is necessary for interpreting the coefficient estimates obtained from the IV approach as Local Average Treatment Effects (LATEs) if there are heterogeneous treatment effects. It requires that any patient who is (not) batched by a sequencer (batcher) would also (not) be batched by a batcher (sequencer) physician. The literature leveraging the judges design typically performs two informal tests for its implications. The first one provides that the first stage should be weakly positive for all subsamples (@dobbie2018effects). The second implication asserts that the instrument constructed by leaving out a particular subsample has predictive power over that same left-out subsample (@bhuller2020incarceration). Appendix Table  [...] presents both of these tests in the two columns for various subsamples of interest.

# Results {#sec:4}

## Reduced-Form Results

In this section, we explore the causal influence of physician batch tendency on patient outcomes and resource utilization in the emergency department. We posit that while batch tendency directly influences the practice of batch ordering tests, both batch tendency and batch ordering are concurrently influenced by a physician's testing inclination. Given that testing inclination directly affects primary outcomes, we include it as a control variable in our regression models to mitigate its confounding effects.

To quantify a physician's testing inclination, we employ a similar approach to that used in measuring physician batch tendency. Specifically, we calculate the physician's residualized leave-out average test rate, which serves as a proxy for their propensity to order tests. It is important to note the strong positive correlation ($r = 0.79$) between batch tendency and testing inclination. This substantial correlation suggests that physicians with a higher propensity to test may also exhibit a higher tendency to batch orders, potentially as a consequence of their testing strategies. Neglecting to account for testing inclination could lead to overestimated effects of batch tendency due to omitted variable bias.

Reduced-form regression results in Table \ref{table:reduced_form} reveal the effect the average total effect of batch tendency on resource utilization and LOS. However, it is important to note that the estimated total effect of batch tendency accounts for both direct and indirect (mediated) effects of batch tendency. We hypothesize that while batch ordering may streamline the testing process, resulting in a quicker completion of a given number of tests, it simultaneously appears to lead to an increase in the total number of tests ordered. The increased testing volume, in turn, is associated with an extended LOS.

\begin{table}[!htbp] 
\centering 
\caption{Reduced Form Estimates for Batch Tendency}
\label{table:reduced_form}
\small
\begin{threeparttable}
\begin{tabular}{lcccccc}
\toprule
 & 72Hr Admission & LOS & Imaging Tests & 72Hr Admission & LOS & Imaging Tests \\
\midrule
Batch Tendency & -0.033 & 2.038** & 4.655*** & -0.063*** & -0.829 & 1.137** \\
 & (0.028) & (0.994) & (0.554) & (0.055) & (2.393) & (0.464) \\
\textit{Baseline Controls} & $\checkmark$ & $\checkmark$ & $\checkmark$ &  $\checkmark$ & $\checkmark$ & $\checkmark$\\
\textit{Testing Inclination} &  &  &  &  $\checkmark$ & $\checkmark$ & $\checkmark$\\
\textit{Any Imaging Ordered} &  &  &  &  $\checkmark$ & $\checkmark$ & $\checkmark$\\
\midrule
N & 15,821 & 15,821 & 15,821 & 15,821 & 15,821 & 15,821 \\
R$^2$ & 0.006 & 0.303 & 0.199 & 0.006 & 0.364 & 0.591 \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Estimates of the reduced form described in the text. Robust standard errors are clustered at the physician level.
\item $^{***} p < 0.001$, $^{**} p < 0.01$, $^{*} p < 0.05$.
\end{tablenotes}
\end{threeparttable}
\end{table}

## Placebo Check

In this section we investigate whether the reduced-form effects observed in Section [...] are due to differences in batch rates across providers or due to other provider differences correlated with batch tendency. We start by studying reduced-form effects among patients with complaints that are never batched, as a ``placebo/falsification check.” By way of example, consider a patient who arrives at the ED with a urinary tact infection—a condition for which patients rarely undergo imaging testing. For such patients, we should expect to see no impact of batch tendency only if high batching and low batching physicians do not systematically differ in other dimensions of care relevant to patient outcomes. Conversely, if we do find a reduced-form effect for these patients, then high batch tendency physicians must systematically differ from low batch tendency physicians in other dimensions of care, beyond batching.

To that end, we restrict attention to ED visits for complaints where batching occurs no more than 5 percent of the time (recall that our baseline sample only includes complaints with a >5 percent batching rate). We estimate a reduced-form regression of each main outcome on physician batch tendency for the subsample, following equation (3). The results of this exercise are displayed in Appendix Table [...]. They show that in contrast to results for our main sample, the association between physician tendency to batch and a given outcome is statistically indistinguishable from zero and much smaller in magnitude for the samples of patients who visit the ED with health conditions that are rarely batched.

## Instrumental Variables Results

To uncover the causal effect of test batching on key emergency department (ED) outcomes, we utilize a two-stage least squares (2SLS) approach, leveraging physician batch-ordering tendencies as an instrument. This strategy isolates the local average treatment effect (LATE) for patients whose diagnostic tests were batched due to their physician’s inclination to batch, rather than patient characteristics or other unobserved factors that may influence outcomes. The IV estimates capture the effect of test batching on ED length of stay, the number of imaging tests ordered, and the likelihood of a return visit within 72 hours with admission.

\begin{table}[!htbp] \centering 
  \caption{IV Results} 
  \label{} 
\begin{tabular}{@{\extracolsep{15pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & Log LOS & Imaging Tests & 72Hr Admission  \\ 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
 Batched & $-$0.504 & 0.692$^{**}$ & $-$0.038 \\ 
  & (1.457) & (0.286) & (0.034) \\ 
  & & & \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular} 
\begin{tablenotes}
\small
\item Estimates of the reduced form described in the text. Robust standard errors are clustered at the physician level.
\item $^{***} p < 0.001$, $^{**} p < 0.01$, $^{*} p < 0.05$.
\end{tablenotes}
\end{table}

*ED Length of Stay*—The IV results in column (1) of Table X indicate that test batching has a negative but statistically insignificant effect on ED length of stay. The coefficient of -0.504 suggests a potential reduction in time spent in the ED, but the confidence interval is wide, indicating that the effect may not be robust across different specifications. This finding implies that while batching may streamline certain diagnostic processes, other operational factors likely play a larger role in determining how long a patient remains in the ED. These results suggest that batching alone does not drastically impact the overall time patients spend in the ED, pointing to the need for further exploration of mechanisms that may moderate this relationship.

*Number of Imaging Tests Ordered*—In column (2), we find a statistically significant and positive effect of test batching on the number of imaging tests ordered. The coefficient of 0.692 (p < 0.05) indicates that batching leads to approximately 69% more imaging tests being conducted during the patient’s ED visit relative to if that patient's tests were sequenced. This result highlights a key consequence of batching: the grouping of diagnostic orders leads to a more intensive use of imaging resources. This increased resource utilization, while contributing to more comprehensive diagnostic evaluations, also raises concerns about the potential for overutilization and the associated costs and patient burden. This effect is particularly relevant in resource-constrained environments, where the trade-off between efficiency and resource use becomes crucial.

*Return Visits with Admission within 72 Hours*—The results in column (3) show that test batching has a negative effect on the likelihood of a return visit with admission within 72 hours, although this effect is not statistically significant. The point estimate of -0.038 suggests a small potential reduction in return visits for patients whose tests were batched, which may reflect more thorough initial diagnostics leading to fewer missed diagnoses or complications. However, the lack of statistical significance means we cannot draw firm conclusions from this result. This finding aligns with previous research suggesting that while batching may improve certain diagnostic processes, its effect on downstream outcomes such as return visits may be more complex and contingent on factors beyond test ordering patterns.

### Mechanism: Mediation Analysis

To further explore the mechanisms through which batching influences ED length of stay (LOS) and return visits with admission within 72 hours, we conducted a formal mediation analysis. Specifically, we assessed whether the number of imaging tests mediates the relationship between batching and these outcomes. This analysis helps to decompose the total effect of batching into direct and indirect effects, with the indirect effect capturing the mediation through imaging tests.

The mediation analysis was performed using a quasi-Bayesian approach with 1,000 simulations, adjusting for relevant covariates, including physician testing inclination, imaging use, day of the week, month of the year, patient complaints, and laboratory performance. The results for both outcomes are summarized in Table \ref{tab:med_results}, which presents the average causal mediation effect (ACME), average direct effect (ADE), and total effect for each outcome.

\begin{table}[h]
\centering
\caption{Mediation Analysis for ED Length of Stay and 72-Hour Return Visits}
\label{tab:med_results}
\begin{tabular}{lcccc}
\hline
Outcome      & Effect Type & Estimate & 95\% CI Lower & 95\% CI Upper \\ \hline
\multirow{4}{*}{\textbf{ED Length of Stay}} 
& & & & \\
    & ACME (Indirect Effect)  & 0.083   & 0.016  & 0.150  \\
    & ADE (Direct Effect)     & -0.550  & -3.686 & 2.300  \\
    & Total Effect            & -0.468  & -3.566 & 2.380  \\ \hline
\multirow{4}{*}{\textbf{72-Hour Return Visits}} 
& & & & \\
    & ACME (Indirect Effect)  & -0.0029  & -0.0059 & 0.000  \\
    & ADE (Direct Effect)     & -0.036   & -0.097  & 0.030  \\
    & Total Effect            & -0.039   & -0.099  & 0.030  \\ \hline
\end{tabular}
\end{table}

\textbf{Incorporating Contemporary Mediation Frameworks.}—In line with the classical literature on mediation (Baron and Kenny, 1986), mediation analysis traditionally proceeds in three steps: (1) testing the effect of the independent variable (batch order) on the dependent variable (LOS/72-hour return) without the mediator (imaging volume); (2) adding the mediator (imaging volume) to the model and testing the direct effect of the independent variable on the dependent variable; and (3) testing the mediator’s effect on the dependent variable in the presence of the independent variable.

However, recent literature has demonstrated that significant results in the zero-order effect (i.e., the total effect) are not required to justify moving forward with mediation analysis. Zhao et al. (2010) and Hayes (2022) argue that requiring significance in each step is an overly restrictive and outdated framework, which often overlooks indirect effects that can be meaningful even when direct effects are not. In our analysis, although we do not always find statistically significant direct effects for batching on LOS or return visits, the mediation pathway through imaging volume remains relevant for understanding the indirect effects of batching.

Thus, following the contemporary guidance in mediation analysis, we proceed with testing both direct and indirect effects simultaneously, allowing us to capture the more nuanced dynamics of how batching influences ED outcomes. This approach provides a richer understanding of the operational impacts of batching, particularly as it pertains to the trade-offs between diagnostic efficiency and resource utilization.

\textbf{Mediation Results for ED Length of Stay.}—The mediation analysis reveals that the indirect effect (ACME) of batching on ED length of stay through the number of imaging tests is $0.083 (p < 0.01)$, suggesting that the increased volume of diagnostic testing marginally extends patient stays in the ED. The direct effect (ADE) of batching on length of stay is negative but not statistically significant ($-0.550, p = 0.71$), and the total effect is also negative ($-0.468$) but insignificant. This indicates that while batching may reduce LOS, the effect is partially offset by the increased testing, leading to no substantial impact on overall ED throughput.

\textbf{Mediation Results for 72-Hour Return Visits.}—For return visits within 72 hours, the indirect effect through the number of imaging tests is negative ($-0.0029, p < 0.05$), indicating that the additional testing associated with batching may slightly reduce the likelihood of a return visit. However, the direct effect ($-0.036$) and total effect ($-0.039$) of batching on return visits are small and not statistically significant, suggesting that while comprehensive diagnostics may reduce return visits, the effect of batching on this outcome is generally weak.

\textbf{Conceptual Framework: Directed Acyclic Graph (DAG).}—Based on the findings from the mediation analysis, we propose a conceptual framework to illustrate the relationships uncovered by our empirical results. The Directed Acyclic Graph (DAG) shown in Figure \ref{fig:dag} provides a visual representation of the causal pathways between batch tendency (Z), batch ordering (X), imaging volume (M), and our outcomes (Y). The graph indicates that the indirect effect of batching operates through the volume of imaging tests, while the direct effect of batching on the outcomes bypasses this mediator.

\begin{figure}[h]
\centering
\includegraphics[width=3in]{model.png}
\caption{Directed Acyclic Graph Representing the Causal Pathways Between Batch Order, Imaging Volume, and ED Outcomes}
\label{fig:dag}
\end{figure}

In this DAG, batch tendency (Z) is used as an instrumental variable to predict batch order (X), which in turn affects both imaging volume (M) and our primary outcomes, including ED length of stay and 72-hour return with admission (Y). The unobserved confounding (U) is assumed to influence both imaging volume and the outcomes directly. The mediation analysis supports this structure, as it confirms that imaging volume mediates the relationship between batch ordering and ED outcomes, albeit modestly. By identifying this framework through our analysis, we provide a data-driven explanation for how batching influences ED operations.

\textbf{Interpretation of Mediation Results.}—These results show that the number of imaging tests serves as a partial mediator in the relationship between batching and ED outcomes. For length of stay, increased testing tends to prolong ED visits, but the overall effect of batching on LOS is negligible. For return visits, the increase in diagnostic testing slightly reduces the likelihood of a return within 72 hours, though the overall effect remains small and insignificant. These findings highlight the trade-offs involved in batching diagnostic orders: while batching may streamline certain processes, the associated increase in testing has modest effects on patient flow and outcomes.

### Time to Result

\newpage

# References